Blangiardo replication

### 2022-09-15 src/example_blangiardo-replication/blangiardo-replication.Rmd
blangiardo-replication html up
Standardise .yml
Iterate blangiardo-replication

---
title: "Blangiardo replication"
author: "Adam Howes"
output: html_document
---

@@

```{r}
y <- c(1.2697, 7.7637, 2.2532, 3.4557, 4.1776, 6.4320, -3.6623, 7.7567, 5.9032, 7.2671,
       -2.3447, 8.0160, 3.5013, 2.8495, 0.6467, 3.2371, 5.8573, -3.3749, 4.1507, 4.3092,
       11.7327, 2.6174, 9.4942, -2.7639, -1.5859, 3.6986, 2.4544, -0.3294, 0.2329, 5.2846)

@@

```{r, fig.height=2, fig.cap="The observations (ordered by index, though this has no interpretation). The prior (mean shown as dashed line) mismatches with the data."}
data.frame(index = 1:30, y = y) %>%
  ggplot(aes(x = index, y = y)) +
  geom_point(shape = 4) +
  labs(title = "The data") +
  geom_hline(yintercept = -3, col = "#666666", lty = "dashed")
```

@@

## `R-INLA`

@@

fit <- inla(formula,
            data = dat,
            control.family = list(hyper = theta_prior),
            control.fixed = list(mean.intercept = x_0, prec.intercept = tau_0))
```

## R implementation

@@

```{r fig.height = 3, fig.cap = "True posterior marginal of theta (blue line) overlaid with the choice of integration points. This naive grid places too many points in regions of the parameter space without much posterior density."}
ggplot(dense_theta) +
  geom_line(aes(x = input, y = output), col = midblue) +
  geom_point(data = blangiardo_theta, aes(x = input, y = output), shape = 4) +
  theme(axis.text.y=element_blank()) +
  labs(title = "Blangiardo integration points for the hyperparameters")
```

@@

One of the INLA vignettes discusses how the user can set their own integration points--see `browseVignettes(package="INLA")`.
The integration points that INLA uses can be found using `fit$joint.hyper`, which we can add on to the above plot:

@@

# Different normalising constant?
ggplot(dense_theta) +
  geom_line(aes(x = input, y = output), col = midblue) +
  geom_point(data = internals_inla_theta, aes(x = input, y = output), shape = 4) +
  theme(axis.text.y=element_blank()) +
  labs(title = "INLA integration points for the hyperparameters") +
  annotate("text", x = 0.175, y = 5e-41,
           label = "Now the points are concentrated \n where there is higher density")
```

@@

## `TMB`

C++ for the negative log joint posterior:

@@

```{r}
h <- MakeADFun(data = dat,
               parameters = param,
               random = c("x"),
               DLL = "blangiardo",
               hessian = TRUE)
```

@@

```{r}
f <- MakeADFun(data = dat,
               parameters = param,
               DLL = "blangiardo",
               hessian = TRUE)
```

@@

tmb_nlminb <- nlminb(start = h$par,
                     objective = h$fn,
                     gradient = h$gr,
                     control = list(iter.max = its, trace = 0))

tmb_optim <- optim(par = h$par,
                   fn = h$fn,
                   gr = h$gn,
                   method = "Brent",
                   lower = -100, # Have to specify upper and lower bounds
                   upper = 100, # when using optimize (1D function)
                   control = list(maxit = its, trace = 0))

@@

```{r}
kable_data <- data.frame(c(exp(fit$internal.summary.hyperpar$mode),
                           exp(r_opt$par),
                           exp(tmb_opt$par)))

>>>>

---
title: "Blangiardo replication"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
    **Background**

    **Task**

    **Findings**

    **Next steps**
---

@@

```{r}
y <- c(
  1.2697, 7.7637, 2.2532, 3.4557, 4.1776, 6.4320, -3.6623, 7.7567, 5.9032, 7.2671,
  -2.3447, 8.0160, 3.5013, 2.8495, 0.6467, 3.2371, 5.8573, -3.3749, 4.1507, 4.3092,
  11.7327, 2.6174, 9.4942, -2.7639, -1.5859, 3.6986, 2.4544, -0.3294, 0.2329, 5.2846
)

@@

```{r, fig.height=2, fig.cap="The observations (ordered by index, though this has no interpretation). The prior (mean shown as dashed line) mismatches with the data."}
data.frame(index = 1:30, y = y) %>%
  ggplot(aes(x = index, y = y)) +
  geom_point(shape = 4) +
  labs(title = "The data") +
  geom_hline(yintercept = -3, col = "#666666", lty = "dashed") +
  theme_minimal()
```

@@

## Implementation using `R-INLA`

This model is simple to fit using the `R-INlA` software directly:

@@

fit <- inla(
  formula,
  data = dat,
  control.family = list(hyper = theta_prior),
  control.fixed = list(mean.intercept = x_0, prec.intercept = tau_0)
)
```

## Implementation using R

@@

```{r fig.height = 3, fig.cap = "True posterior marginal of theta (blue line) overlaid with the choice of integration points. This naive grid places too many points in regions of the parameter space without much posterior density."}
ggplot(dense_theta) +
  geom_line(aes(x = input, y = output), col = midblue) +
  geom_point(data = blangiardo_theta, aes(x = input, y = output), shape = 4) +
  theme(axis.text.y=element_blank()) +
  labs(title = "Blangiardo integration points for the hyperparameters") +
  theme_minimal()
```

@@

One of the INLA vignettes discusses how the user can set their own integration points--see `browseVignettes(package = "INLA")`.
The integration points that INLA uses can be found using `fit$joint.hyper`, which we can add on to the above plot:

@@

# Different normalising constant?
ggplot(dense_theta) +
  geom_line(aes(x = input, y = output), col = midblue) +
  geom_point(data = internals_inla_theta, aes(x = input, y = output), shape = 4) +
  theme(axis.text.y=element_blank()) +
  labs(title = "INLA integration points for the hyperparameters") +
  annotate(
    "text", x = 0.175, y = 5e-41,
    label = "Now the points are concentrated \n where there is higher density"
  ) +
  theme_minimal()
```

@@

## Implementation in `TMB`

C++ for the negative log joint posterior is given by:

@@

```{r}
h <- MakeADFun(
  data = dat,
  parameters = param,
  random = c("x"),
  DLL = "blangiardo",
  hessian = TRUE
)
```

@@

```{r}
f <- MakeADFun(
  data = dat,
  parameters = param,
  DLL = "blangiardo",
  hessian = TRUE
)
```

@@

tmb_nlminb <- nlminb(
  start = h$par,
  objective = h$fn,
  gradient = h$gr,
  control = list(iter.max = its, trace = 0)
)

tmb_optim <- optim(
  par = h$par,
  fn = h$fn,
  gr = h$gn,
  method = "Brent",
  lower = -100, # Have to specify upper and lower bounds
  upper = 100, # when using optimize (1D function)
  control = list(maxit = its, trace = 0)
)

@@

```{r}
#' For some reason the mode in the INLA result below is NA
kable_data <- data.frame(c(exp(fit$internal.summary.hyperpar$mean),
                           exp(r_opt$par),
                           exp(tmb_opt$par)))
